---
title: "Transcript Annotation Database for *Oryza sativa*"
author: "Kabilan S"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Transcript Annotation Database for Oryza sativa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**TxDb.Osativa.MSU.RGAP7** is a Bioconductor-compatible transcript annotation
package for *Oryza sativa* generated from the official **Rice Genome Annotation
Project (RGAP) Release 7** GFF3 annotation based on the **IRGSP-1.0** reference
genome (also known as MSU7).

The package provides genomic coordinates and transcript models, including genes,
transcripts, exons, coding sequences (CDS), and promoter regions. It is intended
for downstream genomic analyses using Bioconductor packages such as
**GenomicFeatures**, **GenomicRanges**, **Signac**, **ArchR**, and **ChIPseeker**.

```{r eval=FALSE}
# Install required Bioconductor packages
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c(
    "GenomicFeatures",
    "GenomicRanges"
))
```

# Loading the package

```{r loading}
library(TxDb.Osativa.MSU.RGAP7)
library(GenomicFeatures)
```

# Package summary

```{r summary}
TxDb.Osativa.MSU.RGAP7
```

# Chromosome information

```{r chromosome}
seqinfo(TxDb.Osativa.MSU.RGAP7)
```

# Available chromosomes

```{r avl_chromosomes}
seqlevels(TxDb.Osativa.MSU.RGAP7)
```
The 12 nuclear chromosomes of the IRGSP‑1.0 / MSU7 assembly. These contain the vast majority of annotated genes. 
`ChrUn` : Unanchored contigs – sequences that could not be placed on any chromosome. Sometimes called “ChrUnknown”.
`ChrSy` : Syngenta pseudomolecule – a separate assembly contributed by Syngenta that contains additional sequence not present in the main IRGSP‑1.0 build. It is often used for the plastid genome or extra scaffolds.

# Gene annotations

Retrieve all annotated genes.

```{r gene_annotations} 
genes(TxDb.Osativa.MSU.RGAP7)
```

# Transcript annotations

```{r transcripts_annotations}
head(transcripts(TxDb.Osativa.MSU.RGAP7))
```

# Exon annotations

```{r exon_annotations}
head(exons(TxDb.Osativa.MSU.RGAP7))
```

# Coding sequence (CDS) annotations

```{r cds}
head(cds(TxDb.Osativa.MSU.RGAP7))
```

# Promoter regions

Extract promoter regions 2 kb upstream and 200 bp downstream of transcription
start sites.

```{r promoters}
head(promoters(
    TxDb.Osativa.MSU.RGAP7,
    upstream = 2000,
    downstream = 200
))
```

# Transcripts grouped by gene

```{r transcripts_gene}
tx_by_gene <- GenomicFeatures::transcriptsBy(TxDb.Osativa.MSU.RGAP7, by = "gene")
tx_by_gene
```

# Exons grouped by transcript

```{r exons_transcript}
exonsBy(TxDb.Osativa.MSU.RGAP7, by = "tx")
```

# Session information

```{r sessioninfo}
sessionInfo()
```
